---
title: "findPC - User Manual"
author: "Haotian Zhuang"
#output: rmarkdown::html_vignette
#vignette: >
#  %\VignetteIndexEntry{UserManual}
#  %\VignetteEngine{knitr::rmarkdown}
#  %\VignetteEncoding{UTF-8}
output:
  html_document:
#    theme: united
#    df_print: kable
date: 'Compiled: `r format(Sys.Date(), "%m/%d/%Y")`'
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction
findPC is a software tool including six methods to automatically select the number of principal components to retain based on the standard deviations explained by each PC. A major advantage of findPC is that the only information required is a numeric vector of standard deviations explained by each PC.

# Installation
findPC software can be installed via Github. Users should have R installed on their computer before installing findPC. R can be downloaded here: http://www.r-project.org/. To install the latest version of findPC package via Github, run following commands in R:
```{r message=FALSE,warning=FALSE}
if (!require("devtools"))
install.packages("devtools")
devtools::install_github("haotian-zhuang/findPC")
library(findPC)
```

# findPC function
The synopsis of findPC is:  
`findPC(sdev,number = 30,method = 'perpendicular line',aggregate = NULL,figure = FALSE)`

The default is to return the number of PCs by Perpendicular line with 30 PCs. The following codes take the 50 PCs of the single-cell RNA-seq data in TSCAN package after preprocessing as an example.
```{r}
data(procdata)
sdev<-prcomp(t(procdata),scale. = T)$sdev[1:50]
head(sdev)
findPC(sdev = sdev)
```

The argument 'sdev' should be sorted in decreasing order.
```{r, error=TRUE}
findPC(sdev = -sdev)
```

## Number
The argument 'number' is a vector including number of PCs used in the following function.
```{r, error=TRUE}
findPC(sdev = sdev,number = 51)
```

```{r}
findPC(sdev = sdev,number = c(24,36,48))
```

## Method
The argument ‘method’ specifies the six methods or returns the six results simultaneously.
```{r, error=TRUE}
findPC(sdev = sdev,method = 'xxx')
```

```{r}
findPC(sdev = sdev,number = c(24,36,48),method = 'all')
```

**Method 1: Piecewise linear model**
```{r}
findPC(sdev = sdev,number = c(24,36,48),method = 'piecewise linear model')
```

**Method 2: First derivative**
```{r}
findPC(sdev = sdev,number = c(24,36,48),method = 'first derivative')
```

**Method 3: Second derivative**
```{r}
findPC(sdev = sdev,number = c(24,36,48),method = 'second derivative')
```

**Method 4: Preceding residual**
```{r}
findPC(sdev = sdev,number = c(24,36,48),method = 'preceding residual')
```

**Method 5: Perpendicular line**
```{r}
findPC(sdev = sdev,number = c(24,36,48),method = 'perpendicular line')
```

**Method 6: K-means clustering**
```{r}
findPC(sdev = sdev,number = c(24,36,48),method = 'k-means clustering')
```

## Aggregate
If users are also interested in the mean, median, or voting (median if all are different, otherwise mode) of the results, the argument ‘aggregate’ will support them.
```{r}
findPC(sdev = sdev,number = c(24,36,48),method = 'all',aggregate = 'mean')
findPC(sdev = sdev,number = c(24,36,48),method = 'all',aggregate = 'median')
findPC(sdev = sdev,number = c(24,36,48),method = 'all',aggregate = 'voting')
```

## Figure
The last argument ‘figure’ provides the option to draw a heatmap showing the results.
```{r, fig.width=8, fig.height=3}
findPC(sdev = sdev,number = c(24,36,48),method = 'all',figure = T)
```

<details>
  <summary>**Session Info**</summary>
```{r}
sessionInfo()
```
</details>
